---
output: html_document
editor_options: 
  chunk_output_type: console
---

# ACLI scenarios

**Gang, see the main page (index.html) for information on how to get the data needed for this segment.**

This .qmd file gets and saves results of ACLI scenarios that were created using the ACLI Excel spreadsheet model:

-   There are about 15 scenarios, each involving changes in model parameters.
-   Each scenario has results for 9 asset classes.
-   Each asset class has 1,000 simulations for 31 years (year 0 plus 30 future years)

The code reads all files for each scenario, stacks them into one long dataframe called `slong`. Each record contains results for one sim for one year, with the following columns:

-   `scenario`: short name for the scenario
-   `fnbase`: short name for the asset class
-   `fdesc`: description of the asset class
-   `sim`: simulation number (1 to 1000)
-   `year`: 0 to 30
-   `asset`: asset value at the end of year (asset is 1 in year 0)
-   `logreturn`: log return for given year
-   `cagr`: compound annual growth rate at end of year

This long dataframe is saved as `acli_scenarios.rds` in the data folder of the project. It's too big to include in the repo so you'll have to create the file by running the code below.

## Setup

```{r}
#| label: setup
#| output: false

source(here::here("report", "_common.R"))

```

```{r}
#| label: constants
#| output: false

# file names and descriptions ----
sfnames <- "
AGGR.csv, Aggressive or specialized equity
BALANCED.csv, Diversified balanced allocation
FIXED.csv, Diversified fixed income
INT.csv, Diversified international equity
SMALL.csv, Intermediate risk equity
US.csv, Diversified large cap US equity
IG_1_5.csv, U.S. Corporate investment grade 1-5 year
IG_Long.csv, U.S. Corporate investment grade 10-30 year
HY.csv, U.S. Corporate high yield (Below investment grade)
"

(fnames <- read_csv(sfnames, col_names = c("fname", "fdesc")))
# create a named vector
(fnamesv <- setNames(fnames$fdesc, fnames$fname |> path_ext_remove() |> str_to_lower()))
# fnamesv["aggr"]


# alternative parameter runs ----
# equity tau -- long-run target volatility
tau_m.05_dir <- "s_tau_m.05" # default minus .05
tau_p.05_dir <- "s_tau_p.05" # default plus .05
tau_p.10_dir <- "s_tau_p.10" # default plus .10

# equity phi -- strength of mean reversion in volatility
phi_m.05_dir <- "s_phi_m.05" # default minus 0.05
phi_p.05_dir <- "s_phi_p.05" # default plus 0.05

phi_m.20_dir <- "s_phi_m.20" # default minus 0.20
phi_p.20_dir <- "s_phi_p.20" # default plus 0.20

# equity sigma(v) -- monthly standard deviation of the log volatility process
sigmav_m.05_dir <- "s_sigmav_m.05" # default minus .05
sigmav_p.05_dir <- "s_sigmav_p.05" # default plus .05

# equity rho -- Correlation to random shock to vol and random return component
rho_m.05_dir <- "s_rho_m.05" # default minus .05
rho_p.05_dir <- "s_rho_p.05" # default plus .05
rho_zero_dir <- "s_rho_zero" # rho set to zero

# interest rate paremeterization directories ----
itau1_m.50_dir <- "s_itau1_m.50" # default minus .50
itau1_p.50_dir <- "s_itau1_p.50" # default plus .50

# parameterizations and descriptions ----
scen_names1 <- "
baseline, Baseline
tau_m.05, Target volatility (tau): reduced by .05 for each equity fund
tau_p.05, Target volatility (tau): increased by .05 for each equity fund
tau_p.10, Target volatility (tau): increased by .10 for each equity fund
phi_m.05, Speed of mean reversion in volatility (phi): reduced by .05 for each equity fund
phi_p.05, Speed of mean reversion in volatility (phi): increased by .05 for each equity fund
phi_m.20, Speed of mean reversion in volatility (phi): reduced by .20 for each equity fund
phi_p.20, Speed of mean reversion in volatility (phi): increased by .20 for each equity fund
sigmav_m.05, Monthly sd of the log volatility process (sigma(v)): reduced by .05 for each equity fund
sigmav_p.05, Monthly sd of the log volatility process (sigma(v)): increased by .05 for each equity fund
rho_m.05, Correlation of random volatility shock vol and random return component (rho): reduced by .05 for each equity fund
rho_p.05, Correlation of random volatility shock and random return component (rho): increased by .05 for each equity fund
rho_zero, Correlation to random shock to vol and random return component (rho) set to zero for each equity fund
itau1_p.50, Mean reversion point for the long rate (tau1): increased by 0.50
itau1_m.50, Mean reversion point for the long rate (tau1): reduced by 0.50
"
(scen_names <- read_csv(scen_names1, col_names = c("scenario", "scendesc")))
saveRDS(scen_names, here::here("data", "acli_scenario_names.rds"))

# rho_zero, Correlation to random shock to vol and random return component (rho) set to zero for each equity fund

```

```{r}
#| label: functions
#| output: false


dirget <- function(dirname, fnames){
  
  # function to read ACLI output data files
  
  ddata <- path(ddir, dirname)
  fndf <- path(ddata, fnames$fname)
  df1 <- vroom(fndf, 
               col_types = cols(.default = col_double()),
               col_names=FALSE, id="fname")
  
  df2 <- df1 |> 
    mutate(scenario=str_remove(dirname, "s_"),
           fname=path_file(fname)) |> 
    left_join(fnames, by = join_by(fname)) |> 
    relocate(scenario) |> 
    lcnames()
  
  df2
}

```

## Get and save ACLI scenarios

Gang, running this next chunk gets all of the ACLI scenarios I created using their spreadsheet model.

```{r}
#| label: get-data
#| output: false

# baseline ----
base <- dirget(base_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)
# count(df1, sim) |> ht()

# equity parameters ----
taum <- dirget(tau_m.05_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

taup <- dirget(tau_p.05_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

taup10 <- dirget(tau_p.10_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

phim05 <- dirget(phi_m.05_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

phip05 <- dirget(phi_p.05_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

phim20 <- dirget(phi_m.20_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

phip20 <- dirget(phi_p.20_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

sigmavm <- dirget(sigmav_m.05_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

sigmavp <- dirget(sigmav_p.05_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

rhom <- dirget(rho_m.05_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

rhop <- dirget(rho_p.05_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

rhoz <- dirget(rho_zero_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

# interest rate parameters ----
itaum <- dirget(itau1_m.50_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

itaup <- dirget(itau1_p.50_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)


stack <- bind_rows(base, 
                   # equity fund parameterizations
                   taum, taup, taup10,
                   phim05, phip05,
                   phim20, phip20,
                   sigmavm, sigmavp,
                   rhom, rhop, rhoz,
                   # interest rate parameterizations
                   itaum, itaup
                   ) |> 
  mutate(fnbase=path_ext_remove(fname) |> str_to_lower())
glimpse(stack)
summary(stack) # make sure there are no NAs or oddball values (neg assets, for example)
count(stack, scenario)
count(stack, fname, fnbase)

# create a  long version of the stacked file, with year starting at 0
slong <- stack |> 
  select(-fname) |> 
  pivot_longer(cols=-c(scenario, fnbase, fdesc, sim),
               values_to = "asset") |> 
  mutate(year=str_sub(name, 2, -1) |> as.integer() - 1) |> 
  select(-name) |> 
  arrange(scenario, fnbase, sim, year) |> 
  mutate(logreturn=log(asset / lag(asset)), # log return, not arithmetic return
         cagr=asset^(1/year) - 1, # compound average growth rate
         .by=scenario, fnbase, sim) |> 
  select(scenario, fnbase, fdesc, sim, year, asset, logreturn, cagr)

glimpse(slong)
saveRDS(slong, here::here("data", "acli_scenarios.rds"))


```

## Explore the results (if you want)

```{r}
#| label: explore
#| output: false


slong <- readRDS(here::here("data", "acli_scenarios.rds"))

slong |> 
  filter(year==30, scenario=="baseline") |> 
  summarise(n=n(), cagrmin=min(cagr), cagrmdn=median(cagr), cagrmax=max(cagr), .by = c(fnbase, fdesc))

# slong |> 
#   filter(year==10, scenario=="baseline", fnbase=="aggr") |> 
#   summarise(uvmeasures(return))
# 
# slong |> 
#   filter(scenario=="baseline", fnbase=="aggr") |> 
#   summarise(uvmeasures(return), .by=year)
# 
# slong |> 
#   filter(scenario=="baseline", year > 0) |> 
#   summarise(uvmeasures(return), .by=fnbase)
# 
# slong |> 
#   filter(fnbase=="aggr", year > 0) |> 
#   summarise(uvmeasures(return), .by=scenario)

# measures <- slong |> 
#   filter(year > 0) |> 
#   summarise(uvmeasures(return), .by=c(scenario, fnbase))
# 
# cagr <- slong |> 
#   filter(year==30) |> 
#   summarise(n=n(), cagrmin=min(cagr), cagrmdn=median(cagr), cagrmax=max(cagr),
#             .by = c(scenario, fnbase, fdesc))
# 
# cagr |> 
#   arrange(desc(cagrmdn)) |> 
#   filter(fnbase=="aggr")


```
