---
output: html_document
editor_options: 
  chunk_output_type: console
---

# ACLI scenarios

**Gang, see the main page (index.html) for information on how to get the data needed for this segment.**

This .qmd file gets and saves results of ACLI scenarios that were created using the ACLI Excel spreadsheet model:

-   There are about 15 scenarios, each involving changes in model parameters.
-   Each scenario has results for 9 asset classes.
-   Each asset class has 1,000 simulations for 31 years (year 0 plus 30 future years)

The code reads all files for each scenario, stacks them into one long dataframe called `slong`. Each record contains results for one sim for one year, with the following columns:

-   `scenario`: short name for the scenario
-   `fnbase`: short name for the asset class
-   `fdesc`: description of the asset class
-   `sim`: simulation number (1 to 1000)
-   `year`: 0 to 30
-   `asset`: asset value at the end of year (asset is 1 in year 0)
-   `logreturn`: log return for given year
-   `cagr`: compound annual growth rate at end of year

This long dataframe is saved as `acli_scenarios.rds` in the data folder of the project. It's too big to include in the repo so you'll have to create the file by running the code below.

## Setup

```{r}
#| label: setup
#| output: false

source(here::here("report", "_common.R"))

```

```{r}
#| label: constants
#| output: false

# file names and descriptions ----
sfnames <- "
AGGR.csv, Aggressive or specialized equity
BALANCED.csv, Diversified balanced allocation
FIXED.csv, Diversified fixed income
INT.csv, Diversified international equity
SMALL.csv, Intermediate risk equity
US.csv, Diversified large cap US equity
IG_1_5.csv, U.S. Corporate investment grade 1-5 year
IG_Long.csv, U.S. Corporate investment grade 10-30 year
HY.csv, U.S. Corporate high yield (Below investment grade)
"

(fnames <- read_csv(sfnames, col_names = c("fname", "fdesc")))
# create a named vector
(fnamesv <- setNames(fnames$fdesc, fnames$fname |> path_ext_remove() |> str_to_lower()))
# fnamesv["aggr"]


# alternative parameter runs ----

parms <- c("phi", "tau", "sigma_v", "rho")
types <- c("lbound", "ubound")
runs <- outer(parms, types, paste, sep = "_") |> as.vector()
(runs <- setNames(runs, runs))

scen_names <- tibble(scenario=runs) |> 
  mutate(temp=str_replace(scenario, "_lbound", "__lbound"),
         temp=str_replace(temp, "_ubound", "__ubound")) |> 
  separate(temp, c("parm", "bound"), sep="__") |> 
  add_row(scenario="baseline", parm="baseline") |> 
  mutate(parmdesc=case_when(
    parm=="baseline" ~ "Baseline",
    parm=="phi" ~ "Speed of mean reversion in volatility (phi)",
    parm=="rho" ~ "Correlation between random shock to vol and random component of return (rho)",
    parm=="sigma_v" ~ "Monthly sd of the log volatility process (sigma(v))",
    parm=="tau" ~ "Target volatility (tau)")) |> 
  mutate(scendesc=case_when(
    bound=="lbound" ~ paste0(parmdesc, ": reduced to lower bound of 90% CI"),
    bound=="ubound" ~ paste0(parmdesc, ": increased to upper bound of 90% CI"),
    scenario=="baseline" ~ "Baseline"))
scen_names

saveRDS(scen_names, here::here("data", "acli_scenario_names.rds"))

# scen_names <- readRDS(here::here("data", "acli_scenario_names.rds"))


# TO COME: interest rate paremeterization directories ----
itau1_m.50_dir <- "s_itau1_m.50" # default minus .50
itau1_p.50_dir <- "s_itau1_p.50" # default plus .50

# rho_zero, Correlation to random shock to vol and random return component (rho) set to zero for each equity fund

```

```{r}
#| label: functions
#| output: false


dirget <- function(dirname, fnames){
  
  # function to read ACLI output data files
  
  ddata <- path(ddir, dirname)
  fndf <- path(ddata, fnames$fname)
  df1 <- vroom(fndf, 
               col_types = cols(.default = col_double()),
               col_names=FALSE, id="fname")
  
  df2 <- df1 |> 
    mutate(scenario=str_remove(dirname, "s_"),
           fname=path_file(fname)) |> 
    left_join(fnames, by = join_by(fname)) |> 
    relocate(scenario) |> 
    lcnames()
  
  df2
}

```

## Get and save ACLI scenarios

Gang, running this next chunk gets all of the ACLI scenarios I created using their spreadsheet model.

```{r}
#| label: get-data
#| output: false

f <- function(dirname, fnames){
  print(dirname)
  dirget(dirname, fnames) |> 
    mutate(sim=row_number(), .by=fname)
}

stack <- purrr::map(scen_names$scenario, \(x) f(x, fnames)) |> 
  list_rbind()|> 
  mutate(fnbase=path_ext_remove(fname) |> str_to_lower())
glimpse(stack)

count(stack, scenario)
count(stack, fnbase, fname, fdesc)
count(stack, sim) |> ht()

# create a  long version of the stacked file, with year starting at 0
slong <- stack |> 
  select(-fname) |> 
  pivot_longer(cols=-c(scenario, fnbase, fdesc, sim),
               values_to = "asset") |> 
  mutate(year=str_sub(name, 2, -1) |> as.integer() - 1) |> 
  select(-name) |> 
  arrange(scenario, fnbase, sim, year) |> 
  mutate(logreturn=log(asset / lag(asset)), # log return, not arithmetic return
         cagr=asset^(1/year) - 1, # compound average growth rate
         .by=scenario, fnbase, sim) |> 
  select(scenario, fnbase, fdesc, sim, year, asset, logreturn, cagr)

glimpse(slong)
saveRDS(slong, here::here("data", "acli_scenarios.rds"))

count(slong, scenario)
count(slong, fnbase, fdesc)

slong |> 
  filter(fnbase=="fixed", year %in% 0:3, sim %in% 1:3) |> 
  arrange(year)


```

## Explore the results (if you want)

```{r}
#| label: explore
#| output: false


slong <- readRDS(here::here("data", "acli_scenarios.rds"))

slong |> 
  filter(year==30, scenario=="baseline") |> 
  summarise(n=n(), cagrmin=min(cagr), cagrmdn=median(cagr), cagrmax=max(cagr), .by = c(fnbase, fdesc))

```
