---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Risk evaluation

Analyze the pension model results previously created.

Summarize the results for each scenario (each row of the result tibble) by unnesting it to allow access to all of the sims and all of the years, calculating summary statistics such as percentiles of the funded ratio or more-sophisticated measures of risk (to come, after Gang's analysis).


## Setup

```{r}
#| label: setup
#| output: false

source(here::here("report", "_common.R"))

slv_scen_names <- read_csv(
"scenario,parm,parmdesc,scendesc
gaussmod, none, none, slv model with gaussian errors
tmod, df, degrees of freedom, slv model with heavy tails
skewmod, alpha, skewness parameter, slv model with skewed returns
levmod, rho, correlation parameter, slv model with volatility correlated with returns") |> 
  mutate(framework="slv")
slv_scen_names

acli_scen_names <- readRDS(here::here("data", "acli_scenario_names.rds"))

scen_names <- bind_rows(
  slv_scen_names,
  acli_scen_names) |> 
  relocate(framework)

```


## Functions to implement Gang's risk measures

Put the functions here for now. Later put them in a separate file.

-   **VaR_loss**: Worst (alpha percentile) asset loss at the end of year t (e.g., 5)


```{r}
#| label: risk-functions
#| output: false

# https://github.com/gchen3/pensionALM/blob/main/alm.R

# Using simulation method to calculate VaR_loss, VaR_erc, VaR_uaal at 95% and 99% over t time period
# Note: Monte Carlo simulation method for VaR: Under the Monte Carlo method, Value at Risk is calculated by randomly 
# creating a number of scenarios for future rates to estimate the change in value for each scenario, 
# and then calculating the VaR according to the worst losses.

VaR_loss <- function(alpha, t, res) {
  # Worst (alpha percentile) asset loss at the end of year t
  res_cut <- res |> filter(year==t)
  assets_tail <- quantile(res_cut$assets, (1-alpha)) # assets at the tail
  assets_initial <- res$assets[1] # first simulation in first year
  
  VaR_loss <-assets_tail - assets_initial
  VaR_loss_per <- VaR_loss / assets_initial # asset % reduction in year t at 1-alpha percentile
  
  return(VaR_loss_per)
  # VaR_loss(alpha=0.95, t=2, res=test)
  }


VaR_erc <- function(alpha, t, res) {
  # ERC amount at year t, alpha percentile
  res_cut <- res |> filter(year==t) # data in year t
  erc_tail <- quantile(res_cut$contrib, alpha) # contribution at alpha percentile
  erc_initial <- res$contrib[1] # initial contribution
  
  VaR_erc_jump <- erc_tail - erc_initial
  VaR_erc_jump_per <-  VaR_erc_jump / erc_initial # percentage jump in erc at alpha percentile in year t

  # revisit whether to return all these values  
  # output <- list(erc_tail=erc_tail, 
  #                VaR_erc_jump=VaR_erc_jump, 
  #                VaR_erc_jump_per=VaR_erc_jump_per)
  
  output <- VaR_erc_jump_per # for now, return a scalar
  
  return(output)
  # VaR_erc(alpha=0.95, t=10, res=test)
}


VaR_uaal <- function(alpha, t, res) {
  # fr in year t, at 1 - alpha percentile 
  res_cut <- res |> filter(year==t)
  VaR_fr <- quantile(res_cut$fr, (1-alpha)) # fr at the 1-alpha percentile
  # just return the VaR_fr since VaR_uaal is easily computable if desired
  return(VaR_fr)
  # VaR_uaal(alpha=0.95, t=10, res=test)
}


prob_fr <- function(fr_threshold, t, res) {
  # Probability of funded ratio falling below x% at the end of t years
  
  res_cut <- res |> filter(year==t)
  n_below <- sum(res_cut$fr < fr_threshold)
  n_percent = n_below / nrow(res_cut)
  return(n_percent)
  # prob_fr(fr_threshold=0.8, t=5, res=test)
 }


prob_erc <- function(erc_threshold, t, res) {
  # Probability of contribution rate higher than x% of payroll at the end of t years
  # See CalPERS "Probability of Employer Contribution Rates Exceeding Given Level (at any point in next 30 years)"
  
  res_cut <- res |> 
    filter(year==t) |> 
    mutate(erc_rate= contrib / payroll)
  
  n_above <- sum(res_cut$erc_rate > erc_threshold)
  n_percent <- n_above / nrow(res_cut)
  return(n_percent)
  # prob_erc(erc_threshold=0.30, t=4, res=test)
}


```


## Get saved pension model results

Get data, show the nested data, and show the head and tail of one scenario unnested.

```{r}
#| label: get-penmod-results
#| output: false

result <- readRDS(here::here("data", "allscen_and_comparison_penmod.rds"))

result

result |> 
  slice_head(n=2) |> 
  select(-lrmat) |> 
  unnest(penmod) |> 
  ht()

result |> 
  select(-lrmat) |> 
  filter(scenario=="baseline") |> 
  unnest(col=penmod) |> 
  ht()

```


## Construct risk measures from pension model results


```{r}
#| label: risk-measures
#| output: false

riskmeasures1 <- result |> 
  # weed out bad results
  filter(!(framework=="acli" & str_detect(scenario, "rho"))) |> # rho doesn't affect ACLI results
  filter(!(framework=="slv" & scenario=="tmod")) |>  # t models did not work well for slv models
  # NOTE: leaves us with about 60 framework-scenario-portfolio-modtype combinations
  select(-c(lrmat, mean, sd)) |> # we don't need these
  unnest(col=penmod) |> # 6.66 million observations
  summarise(mean=first(mean_model),
            sd=first(sd_model),
            vloss_a95t10 = VaR_loss(alpha=0.95, t=10, res=pick(everything())), # asset loss at alpha percentile, year t
            verc_a95t10 = VaR_erc(alpha=0.95, t=10, res=pick(everything())), # ERC amount at year t, alpha percentile
            vuaal_a95t10 = VaR_uaal(alpha=0.95, t=10, res=pick(everything())), # fr in year t, (1 - alpha) percentile 
            probfr_fr70t10 = prob_fr(fr_threshold=0.7, t=10, res=pick(everything())), # prob FR below x% at end year t
            proberc_erc20t10 = prob_erc(erc_threshold=0.20, t=10, res=pick(everything())), # prob ERC above x% payroll at end year t
            mdn_frt10=median(fr[year==10]), # median FR at end year t
            .by=c(framework, scenario, porttype, modtype)) # 60 combos

riskmeasures <- riskmeasures1 |> 
  pivot_longer(-c(framework, scenario, porttype, modtype, mean, sd),
               names_to = "measure") |> # 6 measures for each combo=360 recs
  pivot_wider(names_from = modtype) |> # 3 modtypes means 120 recs
  relocate(measure) |> 
  relocate(model, .after = rnormal) |> 
  mutate(diff=model - rnormal) |> 
  left_join(scen_names |> 
              select(framework, scenario, scendesc), 
            by = join_by(framework, scenario)) |> 
  relocate(scendesc) |> 
  arrange(measure, framework, scenario, porttype)

riskmeasures

riskmeasures |> 
  arrange(porttype, measure, scenario)

riskmeasures |> 
  filter(framework=="acli", scenario=="baseline", porttype=="highequity")

riskmeasures |> 
  filter(scenario=="baseline", porttype=="highfixed")

```


## Model structure risk

To what extent does the choice of models change the model output?

NOTE: maybe add pure asset measure w/o contribs or payouts

### High-equity portfolio

Baseline (Gaussian) SLV model vs. constant returns and model-equivalent random normal returns

<!-- Takeaways: -->

<!-- -   Asset risk, contribution risk, and funded ratio all are worse once we introduce volatility, compared to a constant-returns scenario -->
<!-- -   (Note that when assets fall, contributions may rise, keeping funded ratio relatively stable and also potentially stabilizing assets) -->
<!-- -   Risk measures are slightly worse for the modeled returns than for random normal returns, but the difference is small compared to the difference between random normal and constant. -->


```{r}
#| label: table1-gaussian-highequity
#| eval: false
#| output: false

mod_order <- c("constant", "rnormal", "model")
meas_order <- c("vloss_a95t10", "verc_a95t10", "mdn_frt10")

base_flipped_equity <- riskmeasures |> 
  filter(framework=="slv", scenario=="gaussmod", 
         porttype=="highequity", measure %in% meas_order) |> 
  select(framework, scenario, measure, mean, sd, constant, rnormal, model) |> 
  mutate(measure=factor(measure, levels=meas_order)) |> # to put columns in right order when flipped
  pivot_longer(c(constant, rnormal, model), names_to = "type") |> 
  arrange(measure) |>
  pivot_wider(names_from = measure) |> 
  relocate(type, .after = scenario) |> 
  mutate(type=factor(type, levels=mod_order)) |>
  mutate(sd=ifelse(type=="constant", 0, sd)) |> 
  arrange(type)


base_flipped_equity |> 
  select(-c(framework, scenario, mean, sd)) |> 
  gt() |> 
  tab_header(
    title = html("Selected Risk Measures for Stochastic Volatility Model with Gaussian Errors, High Equity Portfolio"),
    subtitle = html("Normal returns have same mean and SD as SV model")
  ) |>
  cols_label(type = html("Constant,<br>Random normal, or<br>modeled returns"),
             vloss_a95t10 = html("VaR % asset loss<br>Year 10 vs. Year 1<br>(5th percentile)"),
             verc_a95t10 = html("VaR % contribution increase<br>Year 10 vs. Year 1<br>(95th percentile)"),
             mdn_frt10 = html("Median funded ratio<br>Year 10")) |> 
  fmt_percent(columns = c(vloss_a95t10, verc_a95t10),
             decimals = 1) |> 
  fmt_percent(columns = mdn_frt10,
              decimals = 1)
  
# tab_source_note(source_note = "Asset loss at 5th percentile")


#   tab_source_note(source_note = htmltools::HTML('<div style="text-align: left;">Note: This table shows sample data.</div>'))

```

<!-- The next table shows risk measures for selected alternative models: -->

<!-- -   Gaussian errors (as above) -->
<!-- -   Skewed errors -->
<!-- -   Leverage model (volatility correlated with returns) -->


<!-- Takeaways: -->

<!-- -   Asset risk, contribution risk, and funded ratio all are worse once we introduce volatility, compared to a constant-returns scenario -->
<!-- -   (Note that when assets fall, contributions may rise, keeping funded ratio relatively stable and also potentially stabilizing assets) -->
<!-- -   Risk measures are slightly worse for the Gaussian modeled returns than for random normal returns, but the difference is small compared to the difference between random normal and constant. -->
<!-- -   Risk measures are slightly worse for the skewed and leverage models than for the Gaussian model, but again not by a lot compared with the difference between random normal and constant returns -->
<!-- -   The leverage model appears to have the highest risk of the three models shown. It also fit the data best (as measured by AIC). -->


```{r}
#| label: table1-slv-highequity
#| eval: false
#| output: false

mod_order <- c("constant", "rnormal", "gaussmod", "skewmod", "levmod")
mod_labs <- c("constant", "rnormal", "Gaussian", "Skewed", "Leverage")
meas_order <- c("vloss_a95t10", "verc_a95t10", "mdn_frt10")

base_flipped_equity <- riskmeasures |> 
  filter(framework=="slv", 
         porttype=="highequity", measure %in% meas_order) |> 
  select(framework, scenario, measure, mean, sd, constant, rnormal, model) |> 
  mutate(measure=factor(measure, levels=meas_order)) |> # to put columns in right order when flipped
  pivot_longer(c(constant, rnormal, model), names_to = "type") |> 
  filter(!(type %in% c("constant", "rnormal") & scenario != "gaussmod")) |> 
  mutate(type=ifelse(type %in% c("constant", "rnormal"),
                     type,
                     scenario)) |> 
  select(-scenario, -framework) |> 
  arrange(measure) |>
  pivot_wider(names_from = measure) |> 
  relocate(type) |> 
  mutate(type=factor(type, levels=mod_order, labels=mod_labs)) |>
  mutate(sd=ifelse(type=="constant", 0, sd)) |> 
  arrange(type)


base_flipped_equity |> 
  gt() |> 
  tab_header(
    title = html("Selected Risk Measures for Alternative Stochastic Volatility Models, High Equity Portfolio"),
    subtitle = html("Normal returns have same mean and SD as SV model")
  ) |>
  cols_label(type = html("Constant,<br>Random normal, or<br>modeled returns"),
             vloss_a95t10 = html("VaR % asset loss<br>Year 10 vs. Year 1<br>(5th percentile)"),
             verc_a95t10 = html("VaR % contribution increase<br>Year 10 vs. Year 1<br>(95th percentile)"),
             mdn_frt10 = html("Median funded ratio<br>Year 10")) |> 
  fmt_percent(columns = c(vloss_a95t10, verc_a95t10),
             decimals = 1) |> 
  fmt_percent(columns = c(mdn_frt10, mean, sd),
              decimals = 1)
  
# tab_source_note(source_note = "Asset loss at 5th percentile")


#   tab_source_note(source_note = htmltools::HTML('<div style="text-align: left;">Note: This table shows sample data.</div>'))

```

Takeaways:

-   Asset risk, contribution risk, and funded ratio all are worse once we introduce volatility, compared to a constant-returns scenario
-   (Note that when assets fall, contributions may rise, keeping funded ratio relatively stable and also potentially stabilizing assets)
-   Risk measures are slightly worse for the Gaussian modeled returns than for random normal returns, but the difference is small compared to the difference between random normal and constant.
-   Risk measures are slightly worse for the skewed and leverage models than for the Gaussian model, but again not by a lot compared with the difference between random normal and constant returns
-   The leverage model appears to have the highest risk of the three models shown. It also fit the data best (as measured by AIC).

```{r}
#| label: table1-all-highequity
#| eval: true
#| output: true

mod_order <- c("constant", "rnormal", "gaussmod", "skewmod", "levmod", "acli")
mod_labs <- c("constant", "rnormal", "Gaussian", "Skewed", "Leverage", "Complex ESG")
meas_order <- c("vloss_a95t10", "verc_a95t10", "mdn_frt10")

base_flipped_equity <- riskmeasures |> 
  filter(framework=="slv" | (framework=="acli" & scenario=="baseline")) |> 
  filter(porttype=="highequity", measure %in% meas_order) |> 
  select(framework, scenario, measure, mean, sd, constant, rnormal, model) |> 
  mutate(measure=factor(measure, levels=meas_order)) |> # to put columns in right order when flipped
  pivot_longer(c(constant, rnormal, model), names_to = "type") |> 
  filter(!(type %in% c("constant", "rnormal") & scenario != "gaussmod")) |> 
  mutate(type=case_when(
    type %in% c("constant", "rnormal") ~ type,
    framework == "acli" & scenario=="baseline" ~ "acli",
    .default = scenario)) |> 
  select(-scenario, -framework) |> 
  arrange(measure) |>
  pivot_wider(names_from = measure) |> 
  relocate(type) |> 
  mutate(type=factor(type, levels=mod_order, labels=mod_labs)) |>
  mutate(sd=ifelse(type=="constant", 0, sd)) |> 
  arrange(type)


base_flipped_equity |> 
  gt() |> 
  tab_header(
    title = html("Selected Risk Measures for Alternative Stochastic Volatility Models, High Equity Portfolio"),
    subtitle = html("Normal returns have same mean and SD as SV model")
  ) |>
  cols_label(type = html("Constant,<br>Random normal, or<br>modeled returns"),
             vloss_a95t10 = html("VaR % asset loss<br>Year 10 vs. Year 1<br>(5th percentile)"),
             verc_a95t10 = html("VaR % contribution increase<br>Year 10 vs. Year 1<br>(95th percentile)"),
             mdn_frt10 = html("Median funded ratio<br>Year 10")) |> 
  fmt_percent(columns = c(vloss_a95t10, verc_a95t10),
             decimals = 1) |> 
  fmt_percent(columns = c(mdn_frt10, mean, sd),
              decimals = 1) |> 
  tab_footnote(
    footnote = "Complex model differs from other models in multiple ways",
    locations = cells_body(
      columns = c(type),
      rows = type == "Complex ESG"))
  
# tab_source_note(source_note = "Asset loss at 5th percentile")
#   tab_source_note(source_note = htmltools::HTML('<div style="text-align: left;">Note: This table shows sample data.</div>'))

```


```{r}
#| label: equity-plot
#| eval: true
#| output: true

base_flipped_equity |> 
  filter(type != "constant") |> 
  mutate(type=fct_recode(type, "Random\nnormal" = "rnormal")) |> 
  ggplot(aes(vloss_a95t10, verc_a95t10)) +
  geom_point(colour="black") +
  geom_text_repel(aes(label=type), colour="blue") +
  # geom_vline(xintercept = 0) +
  scale_x_continuous(name="Asset % change year 10 vs. year 1, 5th percentile",
                     breaks=seq(-1, 1, by=.01),
                     labels = scales::percent_format(accuracy=1)) +
  scale_y_continuous(name="Contribution % change year 10 vs. year 1, 95th percentile",
                     breaks=seq(-1, 2, by=.01),
                     labels = scales::percent_format(accuracy=1)) +  
  ggtitle("Impact of model structure on risk of extreme asset loss and extreme contribution increase",
          subtitle="High-equity portfolios") +
  labs(caption="All portfolios have mean log return of ~6.25%") +
  theme_bw() +
  theme(plot.title = element_text(size = 11),
        plot.subtitle = element_text(size = 8)) +
  caption_left
  

```


### High fixed portfolio

<!-- Baseline (Gaussian) SLV model vs. constant returns and model-equivalent random normal returns -->

<!-- Takeaways: -->

<!-- -   Asset risk, contribution risk, and funded ratio all are worse once we introduce volatility, compared to a constant-returns scenario -->
<!-- -   (Note that when assets fall, contributions may rise, keeping funded ratio relatively stable and also potentially stabilizing assets) -->
<!-- -   Risk measures are slightly worse for the modeled returns than for random normal returns, but the difference is small compared to the difference between random normal and constant. -->


```{r}
#| label: table1-gaussian-highfixed
#| eval: false
#| output: false

mod_order <- c("constant", "rnormal", "model")
meas_order <- c("vloss_a95t10", "verc_a95t10", "mdn_frt10")

base_flipped_fixed <- riskmeasures |> 
  filter(framework=="slv", scenario=="gaussmod", 
         porttype=="highfixed", measure %in% meas_order) |> 
  select(framework, scenario, measure, mean, sd, constant, rnormal, model) |> 
  mutate(measure=factor(measure, levels=meas_order)) |> # to put columns in right order when flipped
  pivot_longer(c(constant, rnormal, model), names_to = "type") |> 
  arrange(measure) |>
  pivot_wider(names_from = measure) |> 
  relocate(type, .after = scenario) |> 
  mutate(type=factor(type, levels=mod_order)) |>
  mutate(sd=ifelse(type=="constant", 0, sd)) |> 
  arrange(type)


base_flipped_fixed |> 
  select(-c(framework, scenario, mean, sd)) |> 
  gt() |> 
  tab_header(
    title = html("Selected Risk Measures for Stochastic Volatility Model with Gaussian Errors, High Fixed Portfolio"),
    subtitle = html("Normal returns have same mean and SD as SV model")
  ) |>
  cols_label(type = html("Constant,<br>random normal, or<br>modeled returns"),
             vloss_a95t10 = html("VaR % asset loss<br>Year 10 vs. Year 1<br>(5th percentile)"),
             verc_a95t10 = html("VaR % contribution increase<br>Year 10 vs. Year 1<br>(95th percentile)"),
             mdn_frt10 = html("Median funded ratio<br>Year 10")) |> 
  fmt_percent(columns = c(vloss_a95t10, verc_a95t10),
             decimals = 1) |> 
  fmt_percent(columns = mdn_frt10,
              decimals = 1)
  
# tab_source_note(source_note = "Asset loss at 5th percentile")


#   tab_source_note(source_note = htmltools::HTML('<div style="text-align: left;">Note: This table shows sample data.</div>'))

```

<!-- The next table shows risk measures for selected alternative models: -->

<!-- -   Gaussian errors (as above) -->
<!-- -   Skewed errors -->
<!-- -   Leverage model (volatility correlated with returns) -->


<!-- Takeaways: -->

<!-- -   Asset risk, contribution risk, and funded ratio all are worse once we introduce volatility, compared to a constant-returns scenario -->
<!-- -   (Note that when assets fall, contributions may rise, keeping funded ratio relatively stable and also potentially stabilizing assets) -->
<!-- -   Risk measures are slightly worse for the Gaussian modeled returns than for random normal returns, but the difference is small compared to the difference between random normal and constant. -->
<!-- -   Risk measures for the skewed model are slightly better than for the Gaussian model, but again not by a lot compared with the difference between random normal and constant returns -->
<!-- -   Risk measures for the leverage model are moderately worse for the leverage model than for the Gaussian model. -->
<!-- -   As with the high-equity portfolio, the leverage model appears to have the highest risk of the three models shown. It also fit the data best (as measured by AIC). -->


```{r}
#| label: table1-slv-highfixed
#| eval: false
#| output: false

mod_order <- c("constant", "rnormal", "gaussmod", "skewmod", "levmod")
mod_labs <- c("constant", "rnormal", "Gaussian", "Skewed", "Leverage")
meas_order <- c("vloss_a95t10", "verc_a95t10", "mdn_frt10")

base_flipped_fixed <- riskmeasures |> 
  filter(framework=="slv", 
         porttype=="highfixed", measure %in% meas_order) |> 
  select(framework, scenario, measure, mean, sd, constant, rnormal, model) |> 
  mutate(measure=factor(measure, levels=meas_order)) |> # to put columns in right order when flipped
  pivot_longer(c(constant, rnormal, model), names_to = "type") |> 
  filter(!(type %in% c("constant", "rnormal") & scenario != "gaussmod")) |> 
  mutate(type=ifelse(type %in% c("constant", "rnormal"),
                     type,
                     scenario)) |> 
  select(-scenario, -framework) |> 
  arrange(measure) |>
  pivot_wider(names_from = measure) |> 
  relocate(type) |> 
  mutate(type=factor(type, levels=mod_order, labels=mod_labs)) |>
  mutate(sd=ifelse(type=="constant", 0, sd)) |> 
  arrange(type)

base_flipped_fixed |> 
  gt() |> 
  tab_header(
    title = html("Selected Risk Measures for Alternative Stochastic Volatility Models, High Fixed Poftfolio"),
    subtitle = html("Normal returns have same mean and SD as SV model")
  ) |>
  cols_label(type = html("Constant,<br>Random normal, or<br>modeled returns"),
             vloss_a95t10 = html("VaR % asset loss<br>Year 10 vs. Year 1<br>(5th percentile)"),
             verc_a95t10 = html("VaR % contribution increase<br>Year 10 vs. Year 1<br>(95th percentile)"),
             mdn_frt10 = html("Median funded ratio<br>Year 10")) |> 
  fmt_percent(columns = c(vloss_a95t10, verc_a95t10),
             decimals = 1) |> 
  fmt_percent(columns = c(mdn_frt10, mean, sd),
              decimals = 1)
  
# tab_source_note(source_note = "Asset loss at 5th percentile")


#   tab_source_note(source_note = htmltools::HTML('<div style="text-align: left;">Note: This table shows sample data.</div>'))

```

Takeaways:

-   Asset risk, contribution risk, and funded ratio all are worse once we introduce volatility, compared to a constant-returns scenario
-   (Note that when assets fall, contributions may rise, keeping funded ratio relatively stable and also potentially stabilizing assets)
-   Risk measures are slightly worse for the Gaussian modeled returns than for random normal returns, but the difference is small compared to the difference between random normal and constant.
-   Risk measures for the skewed model are slightly better than for the Gaussian model, but again not by a lot compared with the difference between random normal and constant returns
-   Risk measures for the leverage model are moderately worse for the leverage model than for the Gaussian model.
-   As with the high-equity portfolio, the leverage model appears to have the highest risk of the three models shown. It also fit the data best (as measured by AIC).

```{r}
#| label: table1-all-highfixed
#| eval: true
#| output: true

mod_order <- c("constant", "rnormal", "gaussmod", "skewmod", "levmod", "acli")
mod_labs <- c("constant", "rnormal", "Gaussian", "Skewed", "Leverage", "Complex ESG")
meas_order <- c("vloss_a95t10", "verc_a95t10", "mdn_frt10")

base_flipped_fixed <- riskmeasures |> 
  filter(framework=="slv" | (framework=="acli" & scenario=="baseline")) |> 
  filter(porttype=="highfixed", measure %in% meas_order) |> 
  select(framework, scenario, measure, mean, sd, constant, rnormal, model) |> 
  mutate(measure=factor(measure, levels=meas_order)) |> # to put columns in right order when flipped
  pivot_longer(c(constant, rnormal, model), names_to = "type") |> 
  filter(!(type %in% c("constant", "rnormal") & scenario != "gaussmod")) |> 
  mutate(type=case_when(
    type %in% c("constant", "rnormal") ~ type,
    framework == "acli" & scenario=="baseline" ~ "acli",
    .default = scenario)) |> 
  select(-scenario, -framework) |> 
  arrange(measure) |>
  pivot_wider(names_from = measure) |> 
  relocate(type) |> 
  mutate(type=factor(type, levels=mod_order, labels=mod_labs)) |>
  mutate(sd=ifelse(type=="constant", 0, sd)) |> 
  arrange(type)


base_flipped_fixed |> 
  gt() |> 
  tab_header(
    title = html("Selected Risk Measures for Alternative Stochastic Volatility Models, High Fixed Poftfolio"),
    subtitle = html("Normal returns have same mean and SD as SV model")
  ) |>
  cols_label(type = html("Constant,<br>random normal, or<br>modeled returns"),
             vloss_a95t10 = html("VaR % asset loss<br>Year 10 vs. Year 1<br>(5th percentile)"),
             verc_a95t10 = html("VaR % contribution increase<br>Year 10 vs. Year 1<br>(95th percentile)"),
             mdn_frt10 = html("Median funded ratio<br>Year 10")) |> 
  fmt_percent(columns = c(vloss_a95t10, verc_a95t10),
             decimals = 1) |> 
  fmt_percent(columns = c(mdn_frt10, mean, sd),
              decimals = 1) |> 
  tab_footnote(
    footnote = "Complex model differs from other models in multiple ways",
    locations = cells_body(
      columns = c(type),
      rows = type == "Complex ESG"))

#  tab_source_note(source_note = "Complex model differs from other models in multiple ways")
  
# tab_source_note(source_note = "Asset loss at 5th percentile")
#   tab_source_note(source_note = htmltools::HTML('<div style="text-align: left;">Note: This table shows sample data.</div>'))

```



```{r}
#| label: fixed-plot
#| eval: true
#| output: true

base_flipped_fixed |> 
  filter(type != "constant") |> 
  mutate(type=fct_recode(type, "Random\nnormal" = "rnormal")) |> 
  ggplot(aes(vloss_a95t10, verc_a95t10)) +
  geom_point(colour="black") +
  geom_text_repel(aes(label=type), colour="blue") +
  # geom_vline(xintercept = 0) +
  scale_x_continuous(name="Asset % change year 10 vs. year 1, 5th percentile",
                     breaks=seq(-1, 1, by=.01),
                     labels = scales::percent_format(accuracy=1)) +
  scale_y_continuous(name="Contribution % change year 10 vs. year 1, 95th percentile",
                     breaks=seq(-1, 2, by=.01),
                     labels = scales::percent_format(accuracy=1)) +  
  ggtitle("Impact of model structure on risk of extreme asset loss and extreme contribution increase",
          subtitle="High fixed-income portfolios") +
  labs(caption="All portfolios have mean log return of ~5.00%") +
  theme_bw() +
  theme(plot.title = element_text(size = 11),
        plot.subtitle = element_text(size = 8)) +
  caption_left

 # theme(
 #    plot.title = element_text(size = 14, hjust = 0.5),  # Adjust the size and position of the title
 #    plot.margin = margin(t = 20, r = 20, b = 20, l = 20)  # Adjust the plot margins
 #  )

```



## Model parameterization risk

To what extent do changes in parameterization change model outputs?

Same portfolio, Same ACLI model, Same mean, Same SD
5-yr VaR loss 5-fr VaR erc 10-yr FR


```{r}
#| label: model-parameterization-risk
#| output: false

scen_names

scen_order <- scen_names |> 
  filter(framework=="acli") |> 
  arrange(parm, bound)

meas_order <- c("vloss_a95t10", "verc_a95t10", "mdn_frt10")

flipped <- riskmeasures |> 
  filter(framework=="acli", measure %in% meas_order) |> 
  select(porttype, scenario, measure, model) |> 
  mutate(measure=factor(measure, levels=meas_order)) |> # to put columns in right order
  arrange(measure) |>
  pivot_wider(names_from = measure, values_from = model) |> 
  mutate(scenario=factor(scenario, levels=scen_order$scenario)) |> 
  arrange(porttype, scenario) |> 
  left_join(scen_order |> select(scenario, scendesc),
            by = join_by(scenario)) |> 
  relocate(scendesc, .after=scenario)

flipped  

```


```{r}
#| label: show-parameter-risk

flipped |> 
  # select(porttype, type, vloss_a95t10, verc_a95t10, probfr_fr70t10) |> 
  arrange(porttype, scenario) |> 
  gt() |> 
  cols_hide(column=porttype) |> 
  tab_header(
    title = html("Selected Risk Measures for Alternative Parameterizations of an ESG Model"),
    subtitle = html("High equity and high bonds portfolios")
  ) |>
  cols_label(porttype="Portfolio type",
             scenario="Parameterization",
             scendesc="Description",
             # type = html("Random normal<br>or<br>modeled returns"),
             vloss_a95t10 = html("VaR asset loss in 10 years"),
             verc_a95t10 = html("VaR contribution increase in 10 years"),
             mdn_frt10 = html("Median funded ratio in 10 years")) |> 
             # probfr_fr70t10 = html("Probability of funded ratio below 70% in 10 years")) |>
  tab_row_group(
    label = "High equity portfolio",
    rows = porttype == "highequity"
  ) |> 
  tab_row_group(
    label = "High bonds portfolio",
    rows = porttype == "highfixed"
  ) |> 
  fmt_number(columns = c(vloss_a95t10, verc_a95t10),
             decimals = 3) |> 
  fmt_percent(columns = mdn_frt10,
              decimals = 1)
```



<!-- ## OLD Exploratory analysis -->

<!-- Take a look at a couple of simple results before running Gang's risk measures. Here we look at the median funded ratio at 30 years. -->

<!-- ```{r} -->
<!-- #| label: explore-data -->
<!-- #| eval: false -->
<!-- #| output: false -->

<!-- tabdata <- result |>  -->
<!--   filter(!str_detect(scenario, "rho")) |>  -->
<!--   select(scenario, porttype, modtype, nsims, nyears, mean_model, sd_model, penmod) |>  -->
<!--   unnest(col=penmod) |>  -->
<!--   filter(year==30) |>  -->
<!--   summarise( -->
<!--     mean_model=first(mean_model), -->
<!--     sd_model=first(sd_model), -->
<!--     frmdn=median(fr), .by=c(scenario, porttype, modtype)) |>  -->
<!--   pivot_wider(names_from = modtype, values_from = frmdn) |>  -->
<!--   select(porttype, scenario, mean_model, sd_model, constant, rnormal, acli) -->

<!-- # tabdata |>  -->
<!-- #   arrange(porttype, scenario) |>  -->
<!-- #   kable(digits=3) -->

<!-- ``` -->

<!-- ### Year 30 median funded ratio, modeled investment returns, deterministic, and random normal returns -->

<!-- This table compares, for each scenario, the median funded ratio at 30 years using ACLI-modeled investment return to the funded ratio using a random normal distribution with the same mean and standard deviation. It is sorted by the largest difference in median funded ratio between modeled and random normal returns. -->

<!-- Observations: -->

<!-- -   In general, the differences from random normal at 30 years hardly look large - the most is 2 percentage points after 30 years. We can easily see swings of 2 percent or more in a single year. -->

<!-- -   Partly, this is because of automatic adjustments built into pension funding - if returns are low, contributions go up, helping to stabilize the funded ratio. That is why we need to look at contributions also. This suggests there's value in just looking at cumulative returns, without considering contributions, to help understand the pure effects of alternative returns. I probably should do that. -->

<!-- -   Modeled median funded ratio is always slightly higher than the random normal counterpart, even though mean log return and sd of log return is virtually identical for both. -->

<!-- -   The difference is much greater for the high fixed-income portfolio than for the high equity portfolio. In these portfolios, mean returns (and standard deviations) are lower than in the high equity portfolios and so median funded ratios are lower as well, but this doesn't explain why the difference from random normal would be large. -->


<!-- ```{r} -->
<!-- #| label: explore-table -->
<!-- #| output: false -->
<!-- #| eval: false -->

<!-- tabdata |>   -->
<!--   arrange(porttype, scenario) |>  -->
<!--   mutate(diff_vsrnorm=acli - rnormal) |>  -->
<!--   left_join(scen_names |>  -->
<!--               select(scenario, scendesc), by = join_by(scenario)) |>  -->
<!--   relocate(scendesc) |>  -->
<!--   arrange(desc(diff_vsrnorm)) |>  -->
<!--   gt() |>  -->
<!--   tab_header( -->
<!--     title = html("Median funded ratio at 30 years for different scenarios and portfolio types - sorted by difference from normal"), -->
<!--     subtitle = html("Modeled portfolio, deterministic (same mean, sd=0), and random normal (same mean, sd)") -->
<!--   ) |> -->
<!--   tab_spanner(columns = contains("_model"), -->
<!--               label="Portfolio mean and sd of log returns") |> -->
<!--   tab_spanner(columns = c(constant, rnormal, acli, diff_vsrnorm), -->
<!--               label="Median funded ratio at 30 years") |> -->
<!--   cols_label(mean_model="mean", -->
<!--              sd_model="sd", -->
<!--              constant="Deterministic", -->
<!--              rnormal=html("Random<br>normal"), -->
<!--              acli = html("Model"), -->
<!--              diff_vsrnorm = html("Model<br>minus<br>normal")) |> -->
<!--   fmt_percent(columns=-c(scendesc, scenario, porttype), -->
<!--              scale=1, -->
<!--              decimals=1) |> -->
<!--   opt_row_striping() -->

<!-- ``` -->

<!-- ### Year 30 median funded ratio, comparison to baseline for same portfolio type and model type -->

<!-- Modeled investment returns, deterministic, and random normal returns -->

<!-- Here the funded ratio baseline is median FR in year 30 for baseline scenario with random normal returns, for the same portfolio type. -->

<!-- frmdn is the median FR in year 30 for the ACLI model. -->

<!-- frdiff is model minus baseline (random normal). -->

<!-- Probably need to think about a better way to compare to baseline. -->

<!-- ```{r} -->
<!-- #| label: old-stuff -->
<!-- #| eval: false -->
<!-- #| output: false -->

<!-- tabdata <- result |>  -->
<!--   filter(!str_detect(scenario, "rho")) |>  -->
<!--   select(scenario, porttype, modtype, nsims, nyears, mean_model, sd_model, penmod) |>  -->
<!--   unnest(col=penmod) |>  -->
<!--   filter(year==30) |>  -->
<!--   summarise( -->
<!--     mean_model=first(mean_model), -->
<!--     sd_model=first(sd_model), -->
<!--     frmdn=median(fr), .by=c(scenario, porttype, modtype)) |>  -->
<!--   group_by(porttype) |>  -->
<!--   mutate(frbaseline=frmdn[scenario=="baseline" & modtype=="rnormal"], -->
<!--          frdiff=frmdn - frbaseline) |>  -->
<!--   relocate(frbaseline, .after=sd_model) |>  -->
<!--   ungroup() |>  -->
<!--   filter(scenario != "baseline") |>  -->
<!--   arrange(porttype, desc(frdiff)) -->

<!-- tabdata |>  -->
<!--   kable(digits=3) -->

<!-- ``` -->



<!-- ```{r} -->
<!-- #| label: explore-plot -->


<!-- tabdata |>  -->
<!--   ggplot(aes(rnormal, model, colour=porttype, label=scenario)) + -->
<!--   geom_point() + -->
<!--   # geom_text_repel(vjust = -0.5, hjust = 0.5, max.overlaps = 15) + -->
<!--   geom_text(vjust = -0.5, hjust = 0.5) + -->
<!--   geom_abline(slope=1) + -->
<!--   scale_color_manual(values=c("red", "blue")) + -->
<!--   scale_x_continuous(name="Funded ratio with random normal returns corresponding to model (same mean, sd)",  -->
<!--                      breaks = seq(0, 1, .05), -->
<!--                      labels = scales::percent_format(accuracy = .1)) + -->
<!--   scale_y_continuous(name="Funded ratio with ACLI-modeled returns",  -->
<!--                      breaks = seq(0, 1, .05), -->
<!--                      labels = scales::percent_format(accuracy = .1)) + -->
<!--   ggtitle(label = "Median funded ratio at 30 years", -->
<!--           subtitle = "ACLI scenarios and portfolios") + -->
<!--   guides(col= guide_legend(title= "Portfolio\ntype")) + -->
<!--   theme_bw() -->

<!-- ``` -->


<!-- ### Comparison to baseline for the same portfolio type -->


<!-- ```{r} -->
<!-- #| label: explore-baseline-table-data -->
<!-- #| output: false -->

<!-- tabdata2 <- tabdata |>  -->
<!--   filter(!str_detect(scenario, "rho")) |>  -->
<!--   select(-rnormal) |>  -->
<!--   rename(frmdn30=model, mean=mean_logreturn, sd=sd_logreturn) |>  -->
<!--   pivot_longer(cols=c(mean, sd, frmdn30), -->
<!--                names_to = "measure", -->
<!--                values_to = "variant") |>  -->
<!--   group_by(porttype, measure) |>  -->
<!--   mutate(baseline=variant[scenario=="baseline"]) |>  -->
<!--   ungroup() |>  -->
<!--   filter(scenario != "baseline") |>  -->
<!--   relocate(variant, .after = baseline) |>  -->
<!--   mutate(diff=variant - baseline) |>  -->
<!--   pivot_wider(names_from = measure, values_from = c(baseline, variant, diff)) |>  -->
<!--   select(-contains("variant")) |>  -->
<!--   left_join(scen_names, by = join_by(scenario)) |>  -->
<!--   relocate(scendesc) |>  -->
<!--   arrange(porttype, desc(diff_frmdn30)) -->

<!-- ``` -->


<!-- #### High-equity portfolio -->

<!-- ```{r} -->

<!-- tabdata2 |>  -->
<!--   filter(porttype=="highequity") |>  -->
<!--   select(-porttype) |>  -->
<!--   gt() |> -->
<!--   tab_header( -->
<!--     title = html("Median funded ratio at 30 years for different scenarios: High equity portfolio"), -->
<!--     subtitle = html("Comparison to baseline scenario") -->
<!--   ) |> -->
<!--   tab_spanner(columns = contains("baseline"), -->
<!--               label="Baseline results") |> -->
<!--   tab_spanner(columns = contains("diff"), -->
<!--               label="Alternative minus baseline") |> -->
<!--   cols_label(baseline_mean="mean return", -->
<!--              diff_mean="mean return", -->
<!--              baseline_sd="sd return", -->
<!--              diff_sd="sd return", -->
<!--              baseline_frmdn30=html("Funded ratio<br>year 30"), -->
<!--              diff_frmdn30=html("Funded ratio<br>year 30") -->
<!--              ) |> -->
<!--   fmt_percent(columns=-c(scendesc, scenario), -->
<!--              scale=1, -->
<!--              decimals=1) |> -->
<!--   opt_row_striping() -->

<!-- ``` -->


<!-- #### High-fixed portfolio -->

<!-- ```{r} -->

<!-- tabdata2 |>  -->
<!--   filter(porttype=="highfixed") |>  -->
<!--   select(-porttype) |>  -->
<!--   gt() |> -->
<!--   tab_header( -->
<!--     title = html("Median funded ratio at 30 years for different scenarios: High fixed income portfolio"), -->
<!--     subtitle = html("Comparison to baseline scenario") -->
<!--   ) |> -->
<!--   tab_spanner(columns = contains("baseline"), -->
<!--               label="Baseline results") |> -->
<!--   tab_spanner(columns = contains("diff"), -->
<!--               label="Alternative minus baseline") |> -->
<!--   cols_label(baseline_mean="mean return", -->
<!--              diff_mean="mean return", -->
<!--              baseline_sd="sd return", -->
<!--              diff_sd="sd return", -->
<!--              baseline_frmdn30=html("Funded ratio<br>year 30"), -->
<!--              diff_frmdn30=html("Funded ratio<br>year 30") -->
<!--              ) |> -->
<!--   fmt_percent(columns=-c(scendesc, scenario), -->
<!--              scale=1, -->
<!--              decimals=1) |> -->
<!--   opt_row_striping() -->

<!-- ``` -->



<!-- ## Construct risk measures from pension model results -->


<!-- ## OLD Calculate risk measures -->

<!-- ```{r} -->
<!-- #| label: risk-measures-OLD -->
<!-- #| output: false -->

<!-- riskmeasures1 <- result |>  -->
<!--   filter(!str_detect(scenario, "rho")) |>  -->
<!--   select(-c(lrmat, mean, sd)) |>  -->
<!--   unnest(col=penmod) |>  -->
<!--   summarise(mean=first(mean_model), -->
<!--             sd=first(sd_model), -->
<!--             vloss_a95t10 = VaR_loss(alpha=0.95, t=10, res=pick(everything())), -->
<!--             verc_a95t10 = VaR_erc(alpha=0.95, t=10, res=pick(everything())), -->
<!--             vuaal_a95t10 = VaR_uaal(alpha=0.95, t=10, res=pick(everything())), -->
<!--             probfr_fr70t10 = prob_fr(fr_threshold=0.7, t=10, res=pick(everything())), -->
<!--             proberc_erc20t10 = prob_erc(erc_threshold=0.20, t=10, res=pick(everything())), -->
<!--             mdn_frt10=median(fr[year==10]), -->
<!--             .by=c(scenario, porttype, modtype)) -->

<!-- riskmeasures <- riskmeasures1 |>  -->
<!--   pivot_longer(-c(scenario, porttype, modtype, mean, sd), -->
<!--                names_to = "measure") |>  -->
<!--   pivot_wider(names_from = modtype) |>  -->
<!--   relocate(acli, .after = rnormal) |>  -->
<!--   relocate(measure) |>  -->
<!--   mutate(diff=acli - rnormal) |>  -->
<!--   left_join(scen_names |>  -->
<!--               select(scenario, scendesc), by = join_by(scenario)) |>  -->
<!--   relocate(scendesc) |>  -->
<!--   arrange(measure, scenario, porttype) -->

<!-- riskmeasures -->

<!-- riskmeasures |>  -->
<!--   arrange(porttype, measure, scenario) -->

<!-- riskmeasures |>  -->
<!--   filter(scenario=="baseline", porttype=="highequity") -->

<!-- riskmeasures |>  -->
<!--   filter(scenario=="baseline", porttype=="highfixed") -->

<!-- ``` -->

