---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Alternative parameterizations

In the scenarios below, we vary key parameters individually or as a group (for example, varying one specific parameter for each of four equity funds).

```{r}
#| label: setup
#| include: false

source(here::here("r", "libraries.r"))
source(here::here("r", "functions_utilities.r"))

```


```{r}
#| label: constants
#| include: false

ddir <- r"(E:\data\soa\acli)"

base_dir <- "s_baseline"
base_fn <- "baseline_ACLI_v1.46 Alt3 12-2020 100yrs.xlsm"


# alternative parameter runds ----
# equity tau -- long-run target volatility
tau_m.06_dir <- "s_tau_m.06" # default minus .06
# tau_m.06_fn <- "tau_m.06_ACLI_v1.46 Alt3 12-2020 100yrs.xlsm"

tau_p.06_dir <- "s_tau_p.06" # default plus .06
# tau_p.06_fn <- "tau_p.06_ACLI_v1.46 Alt3 12-2020 100yrs.xlsm"

# equity phi -- strength of mean reversion in volatility
phi_m.20_dir <- "s_phi_m.20" # default minus 0.20
phi_p.20_dir <- "s_phi_p.20" # default plus 0.20


# file names and descriptions ---
sfnames <- "
AGGR.csv, Aggressive or specialized equity
BALANCED.csv, Diversified balanced allocation
FIXED.csv, Diversified fixed income
INT.csv, Diversified international equity
SMALL.csv, Intermediate risk equity
US.csv, Diversified large cap US equity
IG_Long.csv, Intermediate govt long
"

(fnames <- read_csv(sfnames, col_names = c("fname", "fdesc")))
# create named vector
(fnamesv <- setNames(fnames$fdesc, fnames$fname |> path_ext_remove() |> str_to_lower()))
# fnamesv["aggr"]


# parameterizations and descriptions ----
scen_names1 <- "
baseline, Baseline
tau_m.06, Target volatility (tau) reduced by .06 for each equity fund
tau_p.06, Target volatility (tau) increased by .06 for each equity fund
phi_m.20, Speed of mean reversion in volatility (phi) reduced by .20 for each equity fund
phi_p.20, Speed of mean reversion in volatility (phi) increased by .20 for each equity fund
"
(scen_names <- read_csv(scen_names1, col_names = c("scenario", "scendesc")))

```

```{r}
#| label: functions
#| include: false

# dirname <- base_dir

# function to read ESG output data files

dirget <- function(dirname, fnames){
  ddata <- path(ddir, dirname)
  fndf <- path(ddata, fnames$fname)
  df1 <- vroom(fndf, 
               col_types = cols(.default = col_double()),
               col_names=FALSE, id="fname")
  
  df2 <- df1 |> 
    mutate(scenario=str_remove(dirname, "s_"),
           fname=path_file(fname)) |> 
    left_join(fnames, by = join_by(fname)) |> 
    relocate(scenario) |> 
    lcnames()
  
  df2
}

```

```{r}
#| label: get-data
#| include: false

base <- dirget(base_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)
# count(df1, sim) |> ht()

taum <- dirget(tau_m.06_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

taup <- dirget(tau_p.06_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

phim <- dirget(phi_m.20_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

phip <- dirget(phi_p.20_dir, fnames) |> 
  mutate(sim=row_number(), .by=fname)

stack <- bind_rows(base, 
                   taum, taup, 
                   phim, phip) |> 
  mutate(fnbase=path_ext_remove(fname) |> str_to_lower())
summary(stack) # make sure there are no NAs or oddball values (neg assets, for example)
count(stack, scenario)
count(stack, fname, fnbase)

# create a  long version of the stacked file, with year starting at 0
slong <- stack |> 
  select(-fname) |> 
  pivot_longer(cols=-c(scenario, fnbase, fdesc, sim)) |> 
  mutate(year=str_sub(name, 2, -1) |> as.integer() - 1) |> 
  select(-name) |> 
  arrange(scenario, fnbase, sim, year)

glimpse(slong)

```


#### Summarize alternative variants
```{r}
#| label: tabdata
#| include: false

tabdata1 <- slong |> 
  filter(year==30) |> 
  summarise(n=n(),
            res=list(quantile(value, na.rm=TRUE)), 
            .by = c(scenario, fnbase, fdesc)) |> 
  left_join(scen_names, by = join_by(scenario))

tabdataw <- tabdata1 |> 
  unnest_wider(col = c(res))

tabdatal <- tabdata1 |> 
  unnest_longer(col = c(res))

```

```{r}
#| label: tab-amounts-show
#| include: true

# define the scenarios to keep - put them in the order we want to see them
# scen_names
scenarios <- c("baseline", "tau_m.06", "tau_p.06", "phi_m.20", "phi_p.20")

# define the funds to keep - put them in the order we want to see them
# fnames
funds <- c("us", "int", "small", "aggr", "balanced")

tabdata <- tabdatal |> 
  filter(scenario %in% scenarios,
         fnbase %in% funds,
         res_id=="50%") |> 
  mutate(scenario=factor(scenario, levels=scenarios),
         fnbase=factor(fnbase, levels=funds)) |> 
  arrange(fnbase) |> 
  select(scenario, scendesc, fnbase, res) |> 
  pivot_wider(names_from = fnbase, values_from = res) |> 
  arrange(scenario)

tabdata |> 
  select(-scenario) |> 
  gt() |> 
  tab_header(
    title = html("Asset values at year 30"),
    subtitle = html("Selected scenarios and assets")
    ) |>
  cols_label(
    scendesc="Scenario",
    us=fnamesv["us"],
    int=fnamesv["int"],
    small=fnamesv["small"],
    aggr=fnamesv["aggr"],
    balanced=fnamesv["balanced"]
    )|> 
  fmt_number(columns=-scendesc,
             decimals=2)|> 
  fmt_currency(columns=-scendesc,
               rows=1,
               decimals=2)

```


```{r}
#| label: tab-indexed-show
#| include: true

# define the scenarios to keep - put them in the order we want to see them
# scen_names
scenarios <- c("baseline", "tau_m.06", "tau_p.06", "phi_m.20", "phi_p.20")

# define the funds to keep - put them in the order we want to see them
# fnames
funds <- c("us", "int", "small", "aggr", "balanced")

tabdata1 <- tabdatal |> 
  filter(scenario %in% scenarios,
         fnbase %in% funds,
         res_id=="50%") |> 
  mutate(scenario=factor(scenario, levels=scenarios),
         fnbase=factor(fnbase, levels=funds)) |> 
  arrange(fnbase) |> 
  select(scenario, scendesc, fnbase, res) |> 
  pivot_wider(names_from = fnbase, values_from = res)

baseline_row <- tabdata1 |> 
  filter(scenario == "baseline")

tabdata2 <- tabdata1 |> 
  mutate(across(-c(scenario, scendesc), ~ . / baseline_row[[cur_column()]] - 1))|> 
  arrange(scenario)

tabdata2 |> 
  select(-scenario) |> 
  gt() |> 
  tab_header(
    title = html("Median asset values at year 30, % above or below each asset's baseline value"),
    subtitle = html("Selected scenarios and assets")
    ) |>
  cols_label(
    scendesc="Scenario",
    us=fnamesv["us"],
    int=fnamesv["int"],
    small=fnamesv["small"],
    aggr=fnamesv["aggr"],
    balanced=fnamesv["balanced"]
    )|> 
  fmt_percent(columns=-scendesc,
             decimals=1)

```



#### Long-run target volatility (tau τ)

The average level the logarithm of volatility is expected to revert toward over the long term.

Default values range from approximately 0.125 for U.S. diversified equity to 0.202 for aggressive equity.

Preliminarily, we examine a scenario that reduces this parameter by .06 for each fund, and a scenario that increases it by .06 for each fund.

```{r}
#| label: tau-data
#| include: false

pdata <- slong |> 
  filter(fnbase=="aggr") |> 
  summarise(a50=p50(value),
            a25=p25(value),
            a75=p75(value),
            .by=c(scenario, fnbase, fdesc, year))

```


```{r}
#| label: show-tau
#| include: true

pdata |> 
  pivot_longer(cols=c(a25, a50, a75)) |> 
  ggplot(aes(year, value, colour=scenario)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(name="Assets", breaks=seq(0, 50, 5)) +
  scale_x_continuous(name="Year", breaks=seq(0, 50, 5)) +
  facet_wrap(~name) +
  ggtitle("Aggressive equity fund, asset values by year",
          subtitle="25th, 50th, and 75th percentiles") +
  theme_bw()

```


## Speed of mean reversion in volatility (phi &phi;)

